\documentclass[11pt,a4paper]{article}
\usepackage{amsmath,a4wide,listings,color}

\usepackage{hyperref}

\lstdefinelanguage{GAMS}
{
	basicstyle=\ttfamily\tiny,
	sensitive = false,
  morekeywords= {Set, Alias, Parameters},%
   sensitive=f,%
   %morecomment=[s]{*}
   keywordstyle = \bfseries
  }

\begin{document}

\lstset{language=GAMS,
%frame=shadowbox, rulesepcolor=\color{blue}
}


\section{The simple monopoly}

We begin with a static monopoly with one technology. The monopolist has the following objective function:

\begin{equation*}
	\max_q\Pi = (\alpha-\beta q)q - cq 
\end{equation*}

subject to

\begin{eqnarray*}
	q - K &\leq& 0\\
	q &\geq& 0 
\end{eqnarray*}
 
We get the following Lagrangian.

\begin{equation*}
	\mathcal{L}(q,\lambda)= (\alpha-\beta q)q - cq -\lambda(q-K)
\end{equation*}

The KKT Conditions are as follows.

\begin{eqnarray*}
	0 \leq -\alpha + 2\beta q + c + \lambda &\bot& q \geq 0\\
	0 \leq -q + K &\bot& \lambda \geq 0
\end{eqnarray*}

Consider the numeric example where $\alpha=100$, $\beta=0.75$ and $c=3$. We compare the results with a quantity restricition of $K=55$, and then without it.

\subsection*{ex$\_$monopoly.gms}

\lstinputlisting{ex_monopoly.gms}

\subsection*{ex$\_$monopoly.lst}

\lstinputlisting{ex_monopoly.lst}

\section{The simple monopoly goes dynamic}


We consider again the monopolist who produces with one technology. Now, we extend the time horizon to two stages, and allow for an investment in $t=0$, which translates into more production capacity in $t=1$.

\begin{equation*}
	\max_{q^0,q^1}\Pi = (\alpha-\beta q^0)q^0 - cq^0 +  (\alpha-\beta q^1)q^1 - cq^1 -FI^0
\end{equation*}

subject to

\begin{eqnarray*}
	q^0 - K^0 &\leq& 0\\
	q^1 - K^1 &\leq& 0\\
        K^1 - K^0 - I^0 &=& 0\\
	q^0,q^1,I^0 &\geq& 0 
\end{eqnarray*}
 
We get the following Lagrangian.

\begin{eqnarray*}
	\mathcal{L}(q^0,q^1,\lambda^0,\lambda^1,\phi^1)&=& (\alpha-\beta q^0)q^0 - cq^0 +  (\alpha-\beta q^1)q^1 - cq^1 -FI^0 \\
        && - \lambda^0(q^0-K^0) - \lambda^1(q^1-K^1) - \phi^1(K^1-K^0-I^0)
\end{eqnarray*}

The KKT Conditions are as follows.

\begin{eqnarray*}
	0 \leq -\alpha + 2\beta q^0 + c + \lambda^0 &\bot& q^0 \geq 0\\
        0 \leq -\alpha + 2\beta q^1 + c + \lambda^1 &\bot& q^1 \geq 0\\
	0 \leq -q^0 + K^0 &\bot& \lambda^0 \geq 0 \\
        0 \leq -q^1 + K^1 &\bot& \lambda^1 \geq 0 \\
        0 = K^1 - K^0 - I^0 &\bot& \phi^1 \quad \mbox{free} \\
        0 \leq -\lambda^1 + \phi^1 &\bot& K^1 \geq 0 \\
        0 \leq F - \phi^1 &\bot& I^0 \geq 0 
\end{eqnarray*}


\subsection*{ex$\_$monopoly\_dyn.gms}

\lstinputlisting{ex_monopoly_dyn.gms}

\subsection*{ex$\_$monopoly\_.lst}

\lstinputlisting{ex_monopoly_dyn.lst}


\section{The simple monopoly goes dynamic (with better notation)}


We consider again the monopolist who produces with one technology. Now, we extend the time horizon to two stages, and allow for an investment in $t=0$, which translates into more production capacity in $t=1$.

\begin{equation*}
	\max_{q^t}\Pi = \sum_t\left[(\alpha-\beta q^t)q^t - cq^t  -FI^t\right]
\end{equation*}

subject to

\begin{eqnarray*}
	q^t - K^t &\leq& 0\quad \forall \,t\\
        K^{t+1} - K^t - I^t &=& 0 \quad \forall \, t>0\\
	q^t,I^t,K^t &\geq& 0 
\end{eqnarray*}
 
We get the following Lagrangian.

\begin{eqnarray*}
	\mathcal{L}(q^t,\lambda^t,\phi^t)&=& \sum_t\left[(\alpha-\beta q^t)q^t - cq^t  -FI^t\right] \\
        && \sum_{t\in T}- \lambda^t(q^t-K^t) - \sum_{t\in (T\backslash 0)}\phi^t(K^{t}-K^{t-1}-I^{t-1})
\end{eqnarray*}

The KKT Conditions are as follows.
 
\begin{eqnarray*}
	0 \leq -\alpha + 2\beta q^t + c + \lambda^t &\bot& q^t \geq 0 \quad\forall\,t\\
	0 \leq -q^t + K^t &\bot& \lambda^t \geq 0 \quad\forall\,t\\
        0 = K^{t} - K^{t-1} - I^{t-1} &\bot& \phi^t \quad \mbox{free}, \quad\forall\,t>0 \\
        0 \leq -\lambda^t + \phi^t &\bot& K^t \geq 0  \quad\forall\,t>0 \\
        0 \leq F - \phi^t &\bot& I^{t-1} \geq 0  \quad\forall\,t>0
\end{eqnarray*}


\subsection*{ex$\_$monopoly\_dyn.gms}

\lstinputlisting{ex_monopoly_dyn.gms}

\subsection*{ex$\_$monopoly\_.lst}

\lstinputlisting{ex_monopoly_dyn.lst}


\section{The dynamic monopoly with three stages}

We consider again the monopolist who produces with one technology. Now, we extend the time horizon to three stages, and allow for an investment in $t=0,1$, which translates into more production capacity in $t=1,2$.

\begin{equation*}
	\max_{q^0,q^1}\Pi = (\alpha-\beta q^0)q^0 - cq^0 +  (\alpha-\beta q^1)q^1 - cq^1 + (\alpha-\beta q^2)q^2 - cq^2 -FI^1
\end{equation*}

subject to

\begin{eqnarray*}
	q^0 - K^0 &\leq& 0\\
	q^1 - K^1 &\leq& 0\\
	q^2 - K^2 &\leq& 0\\
        K^1 - K^0 - I^0 &=& 0\\
        K^2 - K^1 - I^1 &=& 0\\
	q^0,q^1,q^1, I^0, I^1, K^0, K^1, K^2 &\geq& 0 
\end{eqnarray*}
 
We get the following Lagrangian.

\begin{eqnarray*}
	\mathcal{L}(q^0,q^1,q2,\lambda^0,\lambda^1,\lambda^2,\phi^1,\phi^2)&=& (\alpha-\beta q^0)q^0 - cq^0 +  (\alpha-\beta q^1)q^1 - cq^1 +  (\alpha-\beta q^2)q^2 - cq^2-FI^1 \\
        && - \lambda^0(q^0-K^0) - \lambda^1(q^1-K^1) - \lambda^1(q^2-K^2) \\
&& - \phi^1(K^1-K^0-I^0) - \phi^2(K^2-K^1-I^1)
\end{eqnarray*}

The KKT Conditions are as follows.

\begin{eqnarray*}
	0 \leq -\alpha + 2\beta q^0 + c + \lambda^0 &\bot& q^0 \geq 0\\
        0 \leq -\alpha + 2\beta q^1 + c + \lambda^1 &\bot& q^1 \geq 0\\
        0 \leq -\alpha + 2\beta q^2 + c + \lambda^2 &\bot& q^2 \geq 0\\
	0 \leq -q^0 + K^0 &\bot& \lambda^0 \geq 0 \\
        0 \leq -q^1 + K^1 &\bot& \lambda^1 \geq 0 \\
        0 \leq -q^2 + K^2 &\bot& \lambda^2 \geq 0 \\
        0 = K^1 - K^0 - I^0 &\bot& \phi^1 \quad \mbox{free} \\
        0 = K^2 - K^1 - I^1 &\bot& \phi^2 \quad \mbox{free} \\
        0 \leq -\lambda^1 + \phi^1 &\bot& K^1 \geq 0 \\
        0 \leq -\lambda^2 + \phi^2 &\bot& K^2 \geq 0 \\
        0 \leq F - \phi^1 &\bot& I^0 \geq 0 \\
        0 \leq F - \phi^2 &\bot& I^1 \geq 0 
\end{eqnarray*}


\subsection*{ex$\_$monopoly\_dyn\_stages.gms}

\lstinputlisting{ex_monopoly_dyn_3stages.gms}

%\subsection*{ex$\_$monopoly\_.lst}

%\lstinputlisting{ex_monopoly_dyn.lst}


\section{The simple monopoly with several technologies}

We continues with a static monopoly with several technologies. The monopolist has the following objective function:

\begin{equation*}
	\max_{q^j}\Pi = (\alpha-\beta \sum_{j\in M}q_j)\sum_{j\in M}q_j - \sum_{j\in M}c_jq_j 
\end{equation*}

subject to

\begin{eqnarray*}
	q_j - K_j &\leq& 0 \quad \forall\,j\\
	q_j, K_j &\geq& 0 \quad \forall\,j
\end{eqnarray*}
 
We get the following Lagrangian.

\begin{equation*}
	\mathcal{L}(q_j,\lambda_j)= (\alpha-\beta \sum_{j\in M}q_j)\sum_{j\in M}q_j - \sum_{j\in M}c_jq_j) - \sum_{j\in M} \lambda_j(q_j-K_j)
\end{equation*}

The KKT Conditions are as follows.

\begin{eqnarray*}
	0 \leq -\alpha + 2\beta \sum_{j\in M}q_j + c_j + \lambda_j &\bot& q_j \geq 0\quad \forall\,j\\
	0 \leq -q_j + K_j &\bot& \lambda_j \geq 0 \quad \forall\,j
\end{eqnarray*}


\subsection*{ex$\_$monopoly\_tech.gms}

\lstinputlisting{ex_monopoly_tech.gms}

\subsection*{ex$\_$monopoly\_tech.lst}

\lstinputlisting{ex_monopoly_tech.lst}

\section{The dynamic monopoly with several technologies}


We consider again the monopolist who produces with several technologies. Now, we extend the time horizon to two stages, and allow for an investment in $t=0$, which translates into more production capacity in $t=1$.

\begin{equation*}
	\max_{q_j^0,q_j^1}\Pi = (\alpha-\beta \sum_{j\in M}q_j^0)\sum_{j\in M}q_j^0 - \sum_{j\in M}c_jq_j^0 +  (\alpha-\beta \sum_{j\in M}q_j^1)\sum_{j\in M}q_j^1 - \sum_{j\in M}c_jq_j^1 - \sum_{j\in M}F_jI_j^0
\end{equation*}

subject to

\begin{eqnarray*}
	q_j^0 - K_j^0 &\leq& 0 \quad \forall\,j\\
	q_j^1 - K_j^1 &\leq& 0 \quad \forall\,j\\
        K_j^1 - K_j^0 - I_j^0 &=& 0 \quad \forall\,j\\
	q_j^0,q_j^1,I_j^0, K_j^0, K_j^1 &\geq& 0 \quad \forall\,j 
\end{eqnarray*}
 
We get the following Lagrangian.

\begin{eqnarray*}
	\mathcal{L}(q_j^0,q_j^1,\lambda_j^0,\lambda_j^1,\phi_j^1)&=& (\alpha-\beta \sum_{j\in M}q_j^0)\sum_{j\in M}q_j^0 - \sum_{j\in M}c_jq_j^0 \\
        && +  (\alpha-\beta \sum_{j\in M}q_j^1)\sum_{j\in M}q_j^1 - \sum_{j\in M}c_jq_j^1  - \sum_{j\in M}F_jI_j^0\\
        && - \sum_{j\in M}\lambda_j^0(q_j^0-K_j^0) - \sum_{j\in M}\lambda_j^1(q_j^1-K^1) - \sum_{j\in M}\phi_j^1(K_j^1-K_j^0-I_j^0)
\end{eqnarray*}

The KKT Conditions are as follows.

\begin{eqnarray*}
	0 \leq -\alpha + 2\beta \sum_{j\in M}q_j^0 + c_j + \lambda_j^0 &\bot& q_j^0 \geq 0\quad \forall\, j\\
	0 \leq -\alpha + 2\beta \sum_{j\in M}q_j^1 + c_j + \lambda_j^1 &\bot& q_j^1 \geq 0\quad \forall\, j\\
	0 \leq -q_j^0 + K_j^0 &\bot& \lambda_j^0 \geq 0 \quad \forall\, j \\
	0 \leq -q_j^1 + K_j^1 &\bot& \lambda_j^1 \geq 0 \quad \forall\, j \\
        0 = K_j^1 - K_j^0 - I_j^0 &\bot& \phi_j^1 \quad \mbox{free}  \quad \forall\, j\\
        0 \leq -\lambda_j^1 + \phi_j^1 &\bot& K_j^1 \geq 0  \quad \forall\, j\\
        0 \leq F_j - \phi_j^1 &\bot& I_j^0 \geq 0  \quad \forall\, j
\end{eqnarray*}


\subsection*{ex$\_$monopoly\_dyn\_tech.gms}

\lstinputlisting{ex_monopoly_dyn_tech.gms}

\subsection*{ex$\_$monopoly\_dyn\_tech.lst}

\lstinputlisting{ex_monopoly_dyn_tech.lst}


\section{The simple oligopoly}

Let us consider now a static oligopoly with one technology and $i$ different players. Each oligopolist has the following objective function:

\begin{equation*}
	\max_{q_i}\Pi_i = (\alpha-\beta \sum_{i\in N}q_i)q_i - cq_i 
\end{equation*}

subject to

\begin{eqnarray*}
	q_i - K_i &\leq& 0  \quad\forall\, i\\
	q_i,K_i &\geq& 0  \quad\forall\, i
\end{eqnarray*}
 
We get the following Lagrangian.

\begin{equation*}
	\mathcal{L}(q_i,\lambda_i)= (\alpha-\beta\sum_{i\in N} q_i)q_i - cq_i -\lambda_i(q_i-K_i)
\end{equation*}

The KKT Conditions are as follows.

\begin{eqnarray*}
	0 \leq -\alpha + \beta\sum_{i\in N} q_i + \beta q_i + c + \lambda_i &\bot& q_i \geq 0 \quad\forall\, i\\
	0 \leq -q_i + K_i &\bot& \lambda_i \geq 0 \quad\forall\, i
\end{eqnarray*}


\subsection*{ex$\_$oligopoly.gms}

\lstinputlisting{ex_oligopoly.gms}

\subsection*{ex$\_$oligopoly.lst}

\lstinputlisting{ex_oligopoly.lst}

\section{The dynamic oligopoly}


We consider again the oligopoly with $i$ players who produce with one technology. Now, we extend the time horizon to two stages, and allow for an investment in $t=0$, which translates into more production capacity in $t=1$.

\begin{equation*}
	\max_{q_i^0,q_i^1}\Pi_i = (\alpha-\beta \sum_{i\in N}q_i^0)q_i^0 - cq^0 +  (\alpha-\beta \sum_{i\in N}q_i^1)q_i^1 - cq_i^1 -FI_i^0
\end{equation*}

subject to

\begin{eqnarray*}
	q_i^0 - K_i^0 &\leq& 0 \quad\forall\, i\\
	q_i^1 - K_i^1 &\leq& 0 \quad\forall\, i\\
        K_i^1 - K_i^0 - I_i^0 &=& 0 \quad\forall\, i\\
	q_i^0,q_i^1,I_i^0, K_i^0, K_i^1 &\geq& 0 \quad\forall\, i
\end{eqnarray*}
 
We get the following Lagrangian.

\begin{eqnarray*}
	\mathcal{L}(q_i^0,q_i^1,\lambda_i^0,\lambda_i^1,\phi_i^1)&=&  (\alpha-\beta \sum_{i\in N}q_i^0)q_i^0 - cq^0 +  (\alpha-\beta \sum_{i\in N}q_i^1)q_i^1 - cq_i^1 -FI_i^0\\
        && - \lambda_i^0(q_i^0-K_i^0) - \lambda_i^1(q_i^1-K_i^1) - \phi_i^1(K_i^1-K_i^0-I_i^0)
\end{eqnarray*}

The KKT Conditions are as follows.

\begin{eqnarray*}
	0 \leq -\alpha + \beta \sum_{i\in N}q_i^0 + \beta q_i^0+ c + \lambda_i^0 &\bot& q_i^0 \geq 0\quad\forall\, i\\
	0 \leq -\alpha + \beta \sum_{i\in N}q_i^1 + \beta q_i^1+ c + \lambda_i^1 &\bot& q_i^1 \geq 0\quad\forall\, i\\
	0 \leq -q_i^0 + K_i^0 &\bot& \lambda_i^0 \geq 0 \quad\forall\, i \\
        0 \leq -q_i^1 + K_i^1 &\bot& \lambda_i^1 \geq 0 \quad\forall\, i\\
        0 = K_i^1 - K_i^0 - I_i^0 &\bot& \phi_i^1 \quad \mbox{free} \quad\forall\, i \\
        0 \leq -\lambda_i^1 + \phi_i^1 &\bot& K_i^1 \geq 0 \quad\forall\, i\\
        0 \leq F - \phi_i^1 &\bot& I_i^0 \geq 0 \quad\forall\, i
\end{eqnarray*}



\subsection*{ex$\_$oligopoly\_dyn.gms}

\lstinputlisting{ex_oligopoly_dyn.gms}

\subsection*{ex$\_$oligopoly\_dyn.lst}

\lstinputlisting{ex_oligopoly_dyn.lst}


\section{The dynamic oligopoly with several technologies}


We consider the oligopoly with $i$ players who produce with several technologies.

\begin{eqnarray*}
	\max_{q_{i,j}^0,q_{i,j}^1}\Pi_i &=& (\alpha-\beta \sum_{i\in N}\sum_{j\in M}q_{i,j}^0)\sum_{j\in M}q_{i,j}^0 - \sum_{j\in M}c_jq_{i,j}^0 \\
&& +  (\alpha-\beta \sum_{i\in N}\sum_{j\in M}q_{i,j}^1)\sum_{j\in M}q_i^1 - \sum_{j\in M}c_jq_{i,j}^1 - \sum_{j\in M}F_jI_{i,j}^0
\end{eqnarray*}

subject to

\begin{eqnarray*}
	q_{i,j}^0 - K_{i,j}^0 &\leq& 0 \quad\forall\, i,j\\
	q_{i,j}^1 - K_{i,j}^1 &\leq& 0 \quad\forall\, i,j\\
        K_{i,j}^1 - K_{i,j}^0 - I_{i,j}^0 &=& 0 \quad\forall\, i,j\\
	q_{i,j}^0,q_{i,j}^1,I_{i,j}^0, K_{i,j}^0, K_{i,j}^1 &\geq& 0 \quad\forall\, i,j
\end{eqnarray*}
 
We get the following Lagrangian.

\begin{eqnarray*}
	\mathcal{L}(q_i^0,q_i^1,\lambda_i^0,\lambda_i^1,\phi_i^1)&=&  (\alpha-\beta \sum_{i\in N}\sum_{j\in M}q_{i,j}^0)\sum_{j\in M}q_{i,j}^0 - \sum_{j\in M}c_jq_{i,j}^0 \\
&& +  (\alpha-\beta \sum_{i\in N}\sum_{j\in M}q_{i,j}^1)\sum_{j\in M}q_i^1 - \sum_{j\in M}c_jq_{i,j}^1 - \sum_{j\in M}F_jI_{i,j}^0\\
        && - \lambda_{i,j}^0(q_{i,j}^0-K_{i,j}^0) - \lambda_{i,j}^1(q_{i,j}^1-K_{i,j}^1) - \phi_{i,j}^1(K_{i,j}^1-K_{i,j}^0-I_{i,j}^0)
\end{eqnarray*}

The KKT Conditions are as follows.

\begin{eqnarray*}
	0 \leq -\alpha + \beta \sum_{i\in N}\sum_{j\in M}q_{i,j}^0 + \beta \sum_{j\in M}q_{i,j}^0+ c + \lambda_{i,j}^0 &\bot& q_{i,j}^0 \geq 0\quad\forall\, i,j\\
	0 \leq -\alpha + \beta \sum_{i\in N}\sum_{j\in M}q_{i,j}^1 + \beta\sum_{j\in M} q_{i,j}^1+ c + \lambda_{i,j}^1 &\bot& q_{i,j}^1 \geq 0\quad\forall\, i,j\\
	0 \leq -q_{i,j}^0 + K_{i,j}^0 &\bot& \lambda_{i,j}^0 \geq 0 \quad\forall\, i,j \\
        0 \leq -q_{i,j}^1 + K_{i,j}^1 &\bot& \lambda_{i,j}^1 \geq 0 \quad\forall\, i,j\\
        0 = K_{i,j}^1 - K_{i,j}^0 - I_{i,j}^0 &\bot& \phi_{i,j}^1 \quad \mbox{free} \quad\forall\, i,j \\
        0 \leq -\lambda_{i,j}^1 + \phi_{i,j}^1 &\bot& K_{i,j}^1 \geq 0 \quad\forall\, i,j\\
        0 \leq F - \phi_{i,j}^1 &\bot& I_{i,j}^0 \geq 0 \quad\forall\, i,j
\end{eqnarray*}

\subsection*{ex$\_$oligopoly\_dyn\_tech.gms}
\lstinputlisting{ex_oligopoly_dyn_tech.gms}

\subsection*{ex$\_$oligopoly\_dyn\_tech.lst}
\lstinputlisting{ex_oligopoly_dyn_tech.lst}

\end{document}